---
title: "Lab3_Group13"
author: "Prudhvi Pedmallu  (prepu690) , Naveen Gabriel (navga709)"
date: "27 September 2018"
output:
  html_document:
    code_folding: hide
fontsize: 11pt
---
<br>

<br>

<br>


```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(plotly)
library(ggplot2)
library(akima)
library(sf)
```

#Assignment 1

<br>

```{r libraries, message=FALSE, warning=FALSE}
#Create-MAPBOX-TOKEN
Sys.setenv('MAPBOX_TOKEN' = 'pk.eyJ1IjoicHJ1cGU2OTAiLCJhIjoiY2ptZHU1NTNmMWFhdzN3cjE1Z21ndW5tcSJ9.jeQeNQEOlRul1wakn_KbpQ')

#Reading data

mosquito<-read.csv('aegypti_albopictus.csv')
```
#1 Distribution of Mosquitoes in 2004 and 2013

In the year 2004, the mosquito type Aedes albopictus is more in the Mississippi region and the Aedes Aegypti type of mosquito is more in Brazil in the North-east region. In the year 2013, we can see the mosquito type Ades Aegypti opacity is very high in Brazil and in the region Taiwan the mosquito type Aedes albopictus opacity is high. In the analysis of 2004 and 2013, we can see that Brazil in the year 2013 affected by the virus zika because of Aedes Aegyptic.In the map, we can find pre-attentive perception problem and we cannot analyze  the affected regions without zoomed in and in the map some regions contains a complex of data, so it is difficult to analyze  data in a particular region.


#Plot-1_2004

<br>

```{r message=FALSE, warning=FALSE}
year2004 <- mosquito %>% 
  select(VECTOR, X, Y, YEAR, COUNTRY) %>%
  filter(YEAR == 2004)

year2004_mosquito <-  year2004 %>%
  plot_mapbox(x = ~X, y = ~Y,
              split = ~VECTOR, hoverinfo='COUNTRY',
              mode = 'scattermapbox') %>%
  layout(title = '2004',
         font = list(color='white'),
         plot_bgcolor = '#FF8000', paper_bgcolor = '#848484',
         mapbox = list(style = 'dark'),
         legend = list(orientation = 'h',
                       font = list(size = 8)),
         margin = list(l = 25, r = 25,
                       b = 25, t = 25,
                       pad = 2))
year2004_mosquito

```

<br>

###Plot-2_2013

<br>

```{r message=FALSE, warning=FALSE}
year2013 <- mosquito %>%
  select(VECTOR, X,Y, YEAR, COUNTRY) %>%
  filter(YEAR == 2013)

year2013_mosquito <-  year2013 %>%
  plot_mapbox(x = ~X, y = ~Y,
              split = ~VECTOR, hoverinfo='COUNTRY',
              mode = 'scattermapbox') %>%
  layout(title = '2013',
         font = list(color='white'),
         plot_bgcolor = '#ff9900', paper_bgcolor = '#848484',
         mapbox = list(style = 'dark'),
         legend = list(orientation = 'h',
                       font = list(size = 8)),
         margin = list(l = 25, r = 25,
                       b = 25, t = 25,
                       pad = 2))
year2013_mosquito

```

<br>

###2 Chloropeth map
In the map the color scaleing is not good ,for example IND(Z=583) and CAN(Z=1) the color is same.
```{r message=FALSE, warning=FALSE}
z_percountry<- (mosquito %>%
             select(VECTOR,YEAR,COUNTRY,COUNTRY_ID) %>%
             group_by(COUNTRY) %>%
             mutate(Z = n()))[,c(-1,-2)] %>% distinct()

z_percountry%>% plot_geo() %>%
  add_trace(z= ~Z, color = ~Z, colors = 'Blues',name = 'mosquito(Z)',
            locations = ~COUNTRY_ID
  ) %>%
  layout(
    title = "Distribution of Mosquitos",
    geo = list(
      projection = list(type = 'equirectangular')
    )
  )
```

<br>

###3 Equidistant and conic equal area projection

In the  map by using logZ we can reduce the scale and we can find colour difference in scaleing, for example we can see difference between two countries like IND and CAN.In equidistant projection the advantage is  all points of a map are equal distance and easy to visible.The dis-advantage is the regions from north to south appear larger when compare to another map.In conic equal area we can see map distortion and we cannot find the  correct distances between the regions.

(A)
```{r message=FALSE, warning=FALSE}
z_percountry$logZ<-log(z_percountry$Z)

z_percountry%>% plot_geo() %>%
  add_trace(z= ~logZ, color = ~logZ, colors = 'Blues',name = 'mosquito(Z)',
            locations = ~COUNTRY_ID
  ) %>%
  layout(
    title = "Distribution of Mosquitos",
    geo = list(
      projection = list(type = 'equidistant')
    )
  )

```


<br>

(B)
```{r message=FALSE, warning=FALSE}
z_percountry$logZ<-log(z_percountry$Z)

z_percountry%>% plot_geo() %>%
  add_trace(z= ~logZ, color = ~logZ, colors = 'Blues',name = 'mosquito(Z)',
            locations = ~COUNTRY_ID
  ) %>%
  layout(
    title = "Distribution of Mosquitos",
    geo = list(
      projection = list(type = 'conic equal area')
    )
  )

```

<br>

#4 Discretization to Identify regions in Brazil Infected by Mosquitoes

In the map, the south-east region  Brazil is most infected by mosquito type Aedes Aegypti. In the map we split the x & y coordinates into a group of 100, due to this we can find the most infected region of mosquito. But analysis of a particular region is difficult because of grouping the data is overplotted.
 
```{r message=FALSE, warning=FALSE}
p = mosquito%>%filter(COUNTRY_ID=="BRA" & YEAR==2013) %>%
 mutate(X1 = cut_interval(X, n= 100))%>% mutate(Y1=cut_interval(Y,n=100))%>% group_by(VECTOR,X1,Y1)%>%summarise(meanX = mean(X),meanY = mean(Y),count = n())

p%>%plot_mapbox(lat = ~meanY,lon =~meanX, alpha = 0.5,split =~VECTOR,color = ~count) %>% 
  layout(title = 'Distribution of Mosquitos',showlegend=TRUE)

```

 

<br>

<br>

<br>


#Assignment 2

<br>

###1 Pre-processing of  Data
Dataset which include information about mean income and age group and counties needs to be pre processed. The counties are renamed to more readable form. Similarly age group are mentioned as a factor like 18-29, 30-49 and
50-64 which are renamed to Young, Adult and Senior respectively. The dataset is then transformed from long to wide with different age group as column. 
```{r message=FALSE, warning=FALSE}
#Reading file
swedfile <- read.csv("Swedish_Household.csv")
swedimage <- readRDS("Sweden_counties.rds")

#Renaming the levels of Age to more readable form and modifying the columns in table
levels(swedfile$age) <- c("Young", "Adult", "Senior")
colnames(swedfile) <- c("Region", "Age", "Mean_Income")

#Renaming the county to more presentable form..
columnsplit <- strsplit(levels(swedfile$Region), " ")
levels(swedfile$Region) <- sapply(columnsplit, function(x) x[[2]])

#Spreading the data from long to wide based on factors of Age
new_swedfile <- tidyr::spread(swedfile, Age, Mean_Income)

#colnames(new_swedfile)<-c("Region","Yo","Mean_Income")
new_swedfile$Region = as.character(new_swedfile$Region)

#Changing the names of region in the file from csv to match region names in sf file
new_swedfile[new_swedfile$Region == "Örebro", ]$Region <- "Orebro"
new_swedfile[new_swedfile$Region == "Västra", ]$Region <- "Västra Götaland"
```


<br>

###2 Violin Plot of Income vs Age Group

With respect to age group (18-29:Young, 30-49:Adult, 50-64:Senior), 3 violin plots shows the density of mean income in SEK(thousands). As quick comparison reveals that the highest of the mean income from young age group is lower than the lowest of mean income from Adult and Senior Group. The violin plot is similar for Adult and Senior. The range of mean income seems to more for Senior age group followed by adult and then young. 

Furthermore, most of the young age group earn mean income around 300K SEK. Most of the people in adult group earn mean income around 470 while for senior age group, around 500K SEK is earned by the most. For each age group, outlier as shown by black dot. For Young outlier value is 385. For adult, The outliers are value is 542,575 and 649. For Senior, the outlier value is 619 and 683. The red dot shows the median value for each age group.
```{r message=FALSE, warning=FALSE}
#Violin Plot for three Age group showing mean income in SEK
p <- ggplot(swedfile, aes(Age, Mean_Income, fill = Age)) + geom_violin() + geom_boxplot(width=0.2, outlier.color="black", fill="white")

p + stat_summary(fun.y = median,
                 geom = "point",
                 color = "Red") + xlab("\n Age Group ") +
                 ylab("Mean Income in SEK\n")
```

<br>

###3.Surface Plot showing the dependance of Senior on Adult and Junior 

From the surface plot it is evident that the value of Senior mean income increases with the increase in income of Adult and young age group. The relationship is nearly linear. Interpolation finds the best fit for the given data range. How good is the linear regression line depends on the future data points. The linear regression might be a good for this model to predict values between the given range but with the small number of data points, there is a chance that linear regression might over fit the data and may be not be a good predictor for unknownn value.
```{r message=FALSE, warning=FALSE}
#Doing a interpolation to get continous variable and creating a surace plot
interpolated = interp(new_swedfile$Young,
                      new_swedfile$Adult,
                      new_swedfile$Senior,
                      duplicate = "mean")

#Creating a surface plot
plot_ly(
    x =  ~ interpolated$x,
    y =  ~ interpolated$y,
    z =  ~ interpolated$z,
    type = "surface"
   
    
) %>% layout(scene = list(
    xaxis = list(title = 'X',gridcolor="grey",gridwidth=2),
    yaxis = list(title = 'Y',gridcolor="grey",gridwidth=2),
    zaxis = list(title = 'Z',gridcolor="grey",gridwidth=2)),
    title= "Surface plot showing dependence of Z on X and Y\n"

    )
```

<br>

###4. Chloropeth map of counties in Sweden based on mean income of Young and Adult age group

Following 2 chloropeth maps are colored on the basis of income value of Young and Adult in different counties of Sweden. Both the chloropeth map appears to be similar in terms of distribution of income with young and adult group at Stockholm appearing to have more mean income in SEK as highlighted with dark blue colors.
<br>
Even though the distribution for young and age mean income appear to be similar but it not clear to compare the value of mean income of Young vs Adult. Though looking at the color scale would facilitate in comparing the income between young and adult but it might need some effort.
```{r message=FALSE, warning=FALSE}
#Modifying the rownames
rownames(new_swedfile) <- new_swedfile$Region

#Adding Mean values of Young and Adult age group corresponding to the name from
#rds file
swedimage$Young <- new_swedfile[swedimage$NAME_1, "Young"]
swedimage$Adult <- new_swedfile[swedimage$NAME_1, "Adult"]

#Plotting chlorpeth plot of sweden counties based on mean income of Young and Adult respectively.)
p <- plot_ly(width=900) %>% layout(
    plot_bgcolor = "#ebfafa",
    title = "Distribution of Mean Income(SEK thousands) for age group 18-29 <br> in counties of Sweden ",
    margin = list(t = 100, 
                  pad = 10)
) %>%
    add_sf(
        data = swedimage,
        color =  ~ Young,
        colors = c("#66ccff","#000066"),
        split = ~ NAME_1,
        showlegend = FALSE
    ) 
p


p2 <- plot_ly(width=900) %>% layout(
    plot_bgcolor = "#ebfafa",
    title = "Distribution of Mean Income(SEK thousands) for age group 30-49 <br> in counties of Sweden ",
    margin = list(t = 100, 
                  pad = 10)
) %>% add_sf(
    data = swedimage,
    color =  ~ Adult,
    split =  ~ NAME_1,
    colors = c("#66ccff","#000066"),
    showlegend = FALSE
) 
p2
```


<br>

###5. Plotting Linköping coordinate

The coordinates of Linköping is shown as a red dot in the chloropeth map of mean income of young age group.
```{r message=FALSE, warning=FALSE}
#Plotting Linköping coordinate in pre existing map
p  %>% add_markers(
    x =  15.624525,
    #Longitude
    y =  58.409814,
    #Latitude
    hoverinfo = "text",
    text =  "Linköping",
    #City Name to Show
    color = I("#d10404")
    ) %>% layout(title = "Red dot showing the location of Linköping.")

```
<br>

<br>

<br>

# Appendix

### Assignment 1, complete code

```{r app1,eval=FALSE}
library(ggplot2)
library(plotly)
library(akima)
library(sf)



#Creating Token to access MAPBOX
Sys.setenv('MAPBOX_TOKEN' = 'pk.eyJ1IjoicHJ1cGU2OTAiLCJhIjoiY2ptZHU1NTNmMWFhdzN3cjE1Z21ndW5tcSJ9.jeQeNQEOlRul1wakn_KbpQ')

#Reading a file
mosquito<-read.csv('aegypti_albopictus.csv')
#-------------------------------------------------------------------------------------------------------------------
#Ass-1.1
#Plot-1_2004

year2004 <- mosquito %>% 
    select(VECTOR, X, Y, YEAR, COUNTRY) %>%
    filter(YEAR == 2004)

year2004_mosquito <-  year2004 %>%
    plot_mapbox(x = ~X, y = ~Y,
                split = ~VECTOR, hoverinfo='COUNTRY',
                mode = 'scattermapbox') %>%
    layout(title = 'Distribution of Mosquitoes in 2004',
           font = list(color='white'),
           plot_bgcolor = '#ff9900', paper_bgcolor = '#848484',
           mapbox = list(style = 'dark'),
           legend = list(orientation = 'h',
                         font = list(size = 8)),
           margin = list(l = 25, r = 25,
                         b = 25, t = 25,
                         pad = 2))
year2004_mosquito
#--------------------------------------------------------------------------------------------------------------------
#Plot-2_2013

year2013 <- mosquito %>%
    select(VECTOR, X,Y, YEAR, COUNTRY) %>%
    filter(YEAR == 2013)

year2013_mosquito <-  year2013 %>%
    plot_mapbox(x = ~X, y = ~Y,
                split = ~VECTOR, hoverinfo='COUNTRY',
                mode = 'scattermapbox') %>%
    layout(title = 'Distribution of Mosquitoes in 2013',
           font = list(color='white'),
           plot_bgcolor = '#ff9900', paper_bgcolor = '#848484',
           mapbox = list(style = 'dark'),
           legend = list(orientation = 'h',
                         font = list(size = 8)),
           margin = list(l = 25, r = 25,
                         b = 25, t = 25,
                         pad = 2))
year2013_mosquito

#-------------------------------------------------------------------------------------------------------------------
#Ass-1.2

z_percountry<- (mosquito %>%
                    select(VECTOR,YEAR,COUNTRY,COUNTRY_ID) %>%
                    group_by(COUNTRY) %>%
                    mutate(Z = n()))[,c(-1,-2)] %>% distinct()

z_percountry%>% plot_geo() %>%
    add_trace(z= ~Z, color = ~Z, colors = 'Blues',name = 'mosquito(Z)',
              locations = ~COUNTRY_ID
    ) %>%
    layout(
        title = "Mosquitoes by Country(total number)",
        geo = list(
            projection = list(type = 'equirectangular')
        )
    )

#-------------------------------------------------------------------------------------------------------------------
#Ass- 1.3
#(A)
z_percountry$logZ<-log(z_percountry$Z)

z_percountry%>% plot_geo() %>%
    add_trace(z= ~logZ, color = ~logZ, colors = 'Blues',name = 'mosquito(Z)',
              locations = ~COUNTRY_ID
    ) %>%
    layout(
        title = "Mosquitoes by Country(log)",
        geo = list(
            projection = list(type = 'equidistant')
        )
    )
#---------------------------------------------------------------------------------------------------------------------

#(B)

z_percountry$logZ<-log(z_percountry$Z)

z_percountry%>% plot_geo() %>%
    add_trace(z= ~logZ, color = ~logZ, colors = 'Blues',name = 'mosquito(Z)',
              locations = ~COUNTRY_ID
    ) %>%
    layout(
        title = "Mosquitoes by Country(log)",
        geo = list(
            projection = list(type = 'conic equal area')
        )
    )
#----------------------------------------------------------------------------------------------------------------------------
#Ass-1.4

p = mosquito%>%filter(COUNTRY_ID=="BRA" & YEAR==2013)

Brazil<- p %>% mutate(X1 = cut_interval(X, n= 100))%>%
    mutate(Y1=cut_interval(Y,n=100))%>% group_by(X1,Y1)%>%
    summarise(meanx = mean(X),meany = mean(Y),N= n())


Brazil %>% plot_mapbox(
    x = ~ meanx,
    y =  ~ meany,
    mode = 'scattermapbox',
    split = ~ N
) %>%
layout(
    title = 'Distribution of Mosquitoes in Brazil',
    font = list(color = 'white'),
    plot_bgcolor = '#ff9900',
    paper_bgcolor = '#848484',
    mapbox = list(style = 'dark'),
    legend = list(orientation = 'h',
                  font = list(size = 8)),
    margin = list(
        l = 25,
        r = 25,
        b = 25,
        t = 25,
        pad = 0.5
    )
)


```


### Assignment 2, complete code

```{r app2,eval=FALSE}
#Reading file
swedfile <- read.csv("Swedish_Household.csv")
swedimage <- readRDS("Sweden_counties.rds")

#Renaming the levels of Age to more readable form and modifying the columns in table
levels(swedfile$age) <- c("Young", "Adult", "Senior")
colnames(swedfile) <- c("Region", "Age", "Mean_Income")

#Renaming the county to simpler form.
columnsplit <- strsplit(levels(swedfile$Region), " ")
levels(swedfile$Region) <- sapply(columnsplit, function(x)
    x[[2]])

#Spreading the data from long to wide based on factors of Age
new_swedfile <- tidyr::spread(swedfile, Age, Mean_Income)
#colnames(new_swedfile)<-c("Region","Yo","Mean_Income")
new_swedfile$Region = as.character(new_swedfile$Region)



#Changing the names of region in the file from csv to match region names in sf file
new_swedfile[new_swedfile$Region == "Örebro", ]$Region <- "Orebro"
new_swedfile[new_swedfile$Region == "Västra", ]$Region <- "Västra Götaland"
    

#Violin Plot for three Age group showing mean income in SEK
p <-
    ggplot(swedfile, aes(Age, Mean_Income, fill = Age)) +
    geom_violin() + geom_boxplot(width = 0.3,
                                 outlier.color = "black",
                                 fill = "white")
                                     
p + stat_summary(fun.y = median,
                 geom = "point",
                 color = "Red") + xlab("\n Age Group ") +
                 ylab("Mean Income in SEK\n")

#Doing a cubic interpolation to get continous variable and creating a surace plot
interpolated = interp(new_swedfile$Young,
                      new_swedfile$Adult,
                      new_swedfile$Senior,
                      duplicate = "mean")

#Creating a surface plot
plot_ly(
    x =  ~ interpolated$x,
    y =  ~ interpolated$y,
    z =  ~ interpolated$z,
    type = "surface"
   
    
) %>% layout(scene = list(
    xaxis = list(title = 'X',gridcolor="grey",gridwidth=2),
    yaxis = list(title = 'Y',gridcolor="grey",gridwidth=2),
    zaxis = list(title = 'Z',gridcolor="grey",gridwidth=2)),
    title= "Surface plot showing dependence of Z on X and Y\n"

    )



#Modifying the rownames
rownames(new_swedfile) <- new_swedfile$Region

#Adding Mean values of Young and Adult age group corresponding to the name from
#rds file
swedimage$Young <- new_swedfile[swedimage$NAME_1, "Young"]
swedimage$Adult <- new_swedfile[swedimage$NAME_1, "Adult"]

#Plotting chlorpeth plot of sweden counties based on mean income of Young and Adult respectively.)
p <- plot_ly(width=900) %>% layout(
    plot_bgcolor = "#ebfafa",
    title = "Distribution of Mean Income(SEK thousands) for age group 18-29 <br> in counties of Sweden ",
    margin = list(t = 100, 
                  pad = 10)
) %>%
    add_sf(
        data = swedimage,
        color =  ~ Young,
        colors = c("#66ccff","#000066"),
        split = ~ NAME_1,
        showlegend = FALSE
    ) 
p


p2 <- plot_ly(width = 900) %>% layout(
    plot_bgcolor = "#ebfafa",
    title = "Distribution of Mean Income(SEK thousands) for age group 30-49 <br> in counties of Sweden ",
    margin = list(t = 100, 
                  pad = 10)
) %>% add_sf(
    data = swedimage,
    color =  ~ Adult,
    split =  ~ NAME_1,
    colors = c("#66ccff","#000066"),
    showlegend = FALSE
) 
p2



#Plotting Linköping coordinate in pre existing map
p  %>% add_markers(
    x =  15.624525,
    #Longitude
    y =  58.409814,
    #Latitude
    hoverinfo = "text",
    text =  "Linköping", #City Name to Show
    color=I("#d10404")
) %>% layout(title = "Red dot showing the location of Linköping.")


```


